\documentclass[a4paper,12pt]{article}
\usepackage{amsmath, amssymb, graphicx, tikz, longtable, hyperref}
\usepackage[a4paper, margin=2cm]{geometry}
\usepackage{circuitikz}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{tabularx}
\title{HUPUF2X8A - APU (Audio Processing Unit) Technical Documentation}
\author{Hutech - PulseForge Series}
\date{\today}

% Custom command for section title pages
\definecolor{titlepagecolor}{RGB}{0,65,120}

\newcommand\sectionbreak{\clearpage}
\titleformat{\section}[display]
  {%
    \pagecolor{titlepagecolor}\thispagestyle{empty}%
    \normalfont\Huge\bfseries\color{white}\filcenter\titlerule[2pt]\vskip5pt%
  }
  {\thesection}
  {10pt}
  {}
  [{\vskip2pt\titlerule[2pt]\sectionbreak\nopagecolor}]

\begin{document}

% Title Page
\include{sections/0_title-page.tex}
\newpage

\tableofcontents
\newpage

% 1. Introduction
\input{sections/1_introduction.tex}
\newpage

% 2. Architecture Overview
\input{sections/2_architecture-overview.tex}
\newpage

% 3. Functional Modules
\input{sections/3_functional-modules.tex}
\newpage

% 4. Register Map and Configuration
\section{Register Map and Configuration}
\subsection{Global Registers}
Listing and description of global control registers.
\subsection{Per-Channel Registers}
Detailed mapping of registers for each channel type.
\subsection{Access Mechanism}
How to interact with registers.
\subsection{Programming Examples}
Example configurations and setups.

% 5. Clock and Timing
\section{Clock and Timing}
\subsection{Master Clock}
Description of clock source and synchronization.
\subsection{Sampling Rate}
Relation between clock speed and audio output.
\subsection{Latency and Constraints}
Expected processing delays.

% 6. Performance Characteristics
\section{Performance Characteristics}
\subsection{Channel Capacity}
Max simultaneous channels and tone generation.
\subsection{Latency}
Measured delays from input to output.
\subsection{Power Consumption}
Energy efficiency and expected consumption.

% 7. Integration and Usage
\section{Integration and Usage}
\subsection{Wiring Diagrams}
Example wiring for headphones and RCA output.
\subsection{Usage Scenarios}
Step-by-step setup for basic sound generation.
\subsection{Programming Interface}
Methods to program and interact with the APU.

% 8. Debugging and Testing
\section{Debugging and Testing}
\subsection{Built-in Diagnostics}
Test modes and validation procedures.
\subsection{Signal Validation}
Test signals and output expectations.
\subsection{Troubleshooting}
Common issues and their resolutions.

% 9. Appendices
\appendix
\section{Electrical Characteristics}
Power and signal specifications.
\section{Example Code}
Sample scripts and configurations.
\section{Revision History}
Track changes and updates to the document.

\end{document}